\documentclass[11pt,journal]{article}
%\usepackage{hyperref}
%\usepackage[breaklinks]{hyperref}
\usepackage{breakurl}
\usepackage{url}
\usepackage{listings}
\usepackage{courier}
\usepackage{amsmath}
%\ifCLASSOPTIONcompsoc
% IEEE Computer Society needs nocompress option
% requires cite.sty v4.0 or later (November 2003)
\usepackage[nocompress]{cite}

%\else
% normal IEEE
\usepackage{cite}
%\fi

\hyphenation{op-tical net-works semi-conduc-tor}
\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in} 
\addtolength{\textwidth}{1.75in}

\addtolength{\topmargin}{-.875in}
\addtolength{\textheight}{1.75in}

\begin{document}
	\title{Sensor Networks and Mobile Data Comminucation, Assignment 1}
	
	\author{Agata~Borkowska,~UID: 1690550}% <-this % stops a space
		%\protect\\
		%\thanks{}}
	
	% The paper headers



	% IEEEtran.cls defaults to using nonbold math in the Abstract.
	% This preserves the distinction between vectors and scalars. However,
	% if the journal you are submitting to favors bold math in the abstract,
	% then you can use LaTeX's standard command \boldmath at the very start
	% of the abstract to achieve this. Many IEEE journals frown on math
	% in the abstract anyway. In particular, the Computer Society does
	% not want either math or citations to appear in the abstract.
	
	% Note that keywords are not normally used for peerreview papers.
	
	% make the title area
	\maketitle
	
	
	% To allow for easy dual compilation without having to reenter the
	% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
	% not be used in maketitle, but will appear (i.e., to be "transported")
	% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
	% is not selected <OR> if conference mode is selected - because compsoc
	% conference papers position the abstract like regular (non-compsoc)
	% papers do!
	%\IEEEdisplaynotcompsoctitleabstractindextext
	% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
	% compsoc under a non-conference mode.
	
	
	% For peer review papers, you can put extra information on the cover
	% page as needed:
	% \ifCLASSOPTIONpeerreview
	% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
	% \fi
	%
	% For peerreview papers, this IEEEtran command inserts a page break and
	% creates the second title. It will be ignored for other modes.
	%\IEEEpeerreviewmaketitle
	
	
	
	\section{Methods 4 - \texttt{lab1mod1.cc}}
	After moving the second node to the position $(100000000.0, 100000000.0, 100000000.0)$ there is no output, implying that no packets were received.
	
	This is not unexpected, as the node which was initially 1 m away from the sender, is now $100000000 \sqrt{3}$. With the log distance propagation model, at this distance the package is too weak to be picked up. In fact, after some trial and error, we found that putting the second node at $(0.0,1000000.0, 0.0)$ is enough for it not to receive the packages.
	
	It is not due to speed, because the packets travel with the speed of light, so they should cover the distance of 100000000 m in 0.33 s.
	
	\section{Methods 5 - \texttt{lab1mod2.cc}}
	We add a third node at $(1.0, 1.0, 1.0)$, and make it the only sink, wile we make the second node (at $(0.0, 1.0, 0.0)$) the source.
	
	This does not change the output from \texttt{lab1.cc}, since we still send 10 packets, and the logging doesn't say which node it was sent from or received at.
	
	\section{Methods 6 - \texttt{lab1mod3.cc} and \texttt{lab1mod4.cc}}
	Changing the duration of the simulation to 60 seconds doesn't change the output, because we still send only 10 packages, one every 0.5 s, which is done in 5 seconds.
	
	However, stopping the simulation after 2 seconds gives only 4 lines of output, as only 4 packets got sent in that time. 
	
	Note: to change how long the simulation is running, we change the parameter in \texttt{Simulator::Stop()}, which is the delay after which the \texttt{Simulator::Run()} is terminated. In the case of a 60 seconds simulation (and in fact the 10 s one as well), the Simulator stopped running after 5 seconds, as 10 packets were sent and no more events were scheduled. If that didn't happen, i.e. it didn't terminate when it executed all scheduled events, it would just idly hang there for the remainder of the time.
	
	%\IEEEPARstart{}{} 
	

	
	% that's all folks
\end{document}

